--
title: "Group project"
subtitle: "Group Number?"
author: 
  - "FirstName LastName"
  - "FirstName LastName"
  - "FirstName LastName"
  - "FirstName LastName"
  - "FirstName LastName"
date: "`r Sys.Date()`"
output:
  html_document:
    theme: readable
    toc: true
    toc_depth: 4
    toc_float: true
    code_download: false
---

```{r setup, include = FALSE}
options(max.print= 120,
        width = 90,
        tibble.width = 80)
knitr::opts_chunk$set(echo= TRUE,
                      cache=FALSE,
                      prompt=FALSE,
                      tidy="styler",
                      comment=NA,
                      message=FALSE,
                      warning=TRUE)

knitr::opts_knit$set(width=90)
knitr::opts_chunk$set(cache.extra = knitr::rand_seed)
set.seed(42)
```

# Checklist {-}

The submission includes the following.

- [ ] RMD document where it's clear what is the code that corresponds to each question. 
- [ ] Dataset
- [ ] html/PDF document with the following
    - [ ] Numbered questions and answers with text and all the necessary code.
    - [ ] Subtitle indicates the group number 
    - [ ] Name of all group members
    - [ ] Details of specification of the work done by group members (e.g., who found the data, who did the pre-processing, who answered which questions, etc).
    - [ ] Statement of technology. Did you use any AI tools? How?


# Group project {-}

For the project, we use the following packages:

```{r, message = FALSE}
## ...
```

## 1. Dataset Selection (0.5pt)
Select a dataset with clusters such as schools, regions, or people with multiple observations per individual. (From for example, https://www.kaggle.com/) It would be a good idea to choose a smallish dataset (not too many rows, e.g., less than 1000) or subset it so that fitting the models doesn't take too long. 

a. Describe the dataset with a couple of short sentences. What was its intended use? Are there papers that reference it? Provide information on how to obtain the dataset, including its source and any necessary preprocessing steps/feature engineering.


<!-- DESCRIBE IT BELOW -->

```{r get-data-direct, echo=TRUE, message=FALSE}
# 0) Install & load httr for HTTP requests
if (!requireNamespace("httr", quietly = TRUE)) {
  install.packages("httr", repos = "https://cran.rstudio.com/")
}
library(httr)

# 1) Kaggle credentials
user <- "barakazor"
key  <- "6bfceda688e1452d09191636ad94eeeb"

# 2) Define working and data directories
wd        <- getwd()
data_dir  <- file.path(wd, "data")
input_dir <- file.path(wd, "Input_Data")
if (!dir.exists(data_dir))  dir.create(data_dir, recursive = TRUE)
if (!dir.exists(input_dir)) dir.create(input_dir, recursive = TRUE)

# 3) Download the ZIP via Kaggle REST API
dataset_slug <- "kirtandelwadia/corporate-credit-rating-with-financial-ratios"
zip_path     <- file.path(data_dir, "credit_rating.zip")
download_url <- paste0("https://www.kaggle.com/api/v1/datasets/download/", dataset_slug)

resp <- GET(
  url             = download_url,
  authenticate(user, key),
  write_disk(zip_path, overwrite = TRUE),
  config(followlocation = TRUE)
)
stop_for_status(resp)

# 4) Unzip into data_dir
unzipped <- unzip(zip_path, exdir = data_dir)
csvs     <- unzipped[grepl("\\.csv$", unzipped)]
if (length(csvs) == 0) stop("No CSV found in the downloaded ZIP.")

# 5) Read the first CSV into R
df_raw <- read.csv(csvs[1], stringsAsFactors = FALSE)

# 6) Save to Input_Data/
write.csv(
  df_raw,
  file      = file.path(input_dir, "df_raw.csv"),
  row.names = FALSE
)
saveRDS(
  df_raw,
  file = file.path(input_dir, "df_raw.rds")
)
```


```{r preprocess-2016, echo=TRUE, message=FALSE}
# Load necessary packages (install if needed)
library(readr)
library(dplyr)
library(janitor)
library(lubridate)

# Read raw data and clean column names explicitly
input_dir <- file.path(wd, "Input_Data")   # adjust as needed
df_raw <- read_csv(file.path(input_dir, "df_raw.csv"))

# Use janitor::clean_names to ensure availability
df_raw <- janitor::clean_names(df_raw)

# Define ordered rating levels
rating_levels <- c("AAA","AA+","AA","AA-","A+","A","A-",
                   "BBB+","BBB","BBB-","BB+","BB","BB-",
                   "B+","B","B-","CCC","CC","C","D")

# Create df with fiscal_year and scaled predictors
df <- df_raw %>%
  mutate(
    rating_ord  = factor(rating, levels = rating_levels, ordered = TRUE),
    sector      = forcats::fct_lump_n(sector, 10),
    corporation = as.factor(corporation),
    fiscal_year = lubridate::year(lubridate::ymd(rating_date))
  ) %>%
  mutate(across(
    c(current_ratio, debt_equity_ratio, gross_margin,
      operating_margin, net_profit_margin, roa_return_on_assets),
    ~ as.numeric(scale(.x)), .names = "z_{col}")) %>%
  tidyr::drop_na(rating_ord, sector, corporation)



# 1. Filter for rating date in 2016 and select only the needed variables
#    - binary_rating: Binary.Rating as DV
#    - rating      : Rating as DV
#    - current_ratio, long_term_debt_capital, debt_equity_ratio,
#      ebitda_margin, operating_cash_flow_per_share, sector: predictors

df_filtered <- df %>%
  filter(fiscal_year == "2016") %>%
  select(
    binary_rating,
    rating,
    current_ratio,
    long_term_debt_capital,
    debt_equity_ratio,
    ebitda_margin,
    operating_cash_flow_per_share,
    sector
  )

# 2. Create two dataframes with different target variables
#    a) df_binary: Binary.Rating as the first column
#    b) df_rating: Rating as the first column

df_binary <- df_filtered %>%
  select(
    Binary.Rating = binary_rating,
    Current.Ratio                   = current_ratio,
    Long.term.Debt...Capital        = long_term_debt_capital,
    Debt.Equity.Ratio               = debt_equity_ratio,
    EBITDA.Margin                   = ebitda_margin,
    Operating.Cash.Flow.Per.Share   = operating_cash_flow_per_share,
    Sector                          = sector
  )


df_rating <- df_filtered %>%
  select(
    Rating                          = rating,
    Current.Ratio                   = current_ratio,
    Long.term.Debt...Capital        = long_term_debt_capital,
    Debt.Equity.Ratio               = debt_equity_ratio,
    EBITDA.Margin                   = ebitda_margin,
    Operating.Cash.Flow.Per.Share   = operating_cash_flow_per_share,
    Sector                          = sector
  )
```


b. Report the number of observations, columns (with their meaning) and their data types. Indicate clearly what you will use as dependent variable/label. 

<!-- REPORT IT BELOW -->

## 2. Split the data and tranform columns as necessary. (0.5pt)
 Split the data into training (80%) and test set (80%). Transform the columns if necessary.

```{r splitdata, echo=TRUE, message=FALSE}
# Load required package for splitting
# (dplyr already loaded)

set.seed(123) # for reproducibility

# Split df_binary into 80% train and 20% test
n_binary <- nrow(df_binary)
train_idx_binary <- sample(seq_len(n_binary), size = 0.8 * n_binary)
train_binary <- df_binary[train_idx_binary, ]
test_binary <- df_binary[-train_idx_binary, ]

# Ensure factor levels of Binary.Rating are consistent
train_binary$Binary.Rating <- factor(train_binary$Binary.Rating)
test_binary$Binary.Rating <- factor(test_binary$Binary.Rating, levels = levels(train_binary$Binary.Rating))

# Split df_rating into 80% train and 20% test
n_rating <- nrow(df_rating)
train_idx_rating <- sample(seq_len(n_rating), size = 0.8 * n_rating)
train_rating <- df_rating[train_idx_rating, ]
test_rating <- df_rating[-train_idx_rating, ]

# Ensure factor levels of Rating are consistent
train_rating$Rating <- factor(train_rating$Rating)
test_rating$Rating <- factor(test_rating$Rating, levels = levels(train_rating$Rating))
```

## 3. Model Exploration (3pt)

a. Fit multiple appropriate models to the dataset (as many models as there are members in the group, with a minimum of two models). Models might vary in the multilevel structure, informativeness of their priors (but not just trivial changes), model of the data/likelihood, etc. (I recommend not to use no pooling models since they tend to take a long time and it's very hard to assign good priors). **Assess if models converged: Report how the traceplots looks like, highest Rhat, number of effective samples, etc. If didn't converge, address the issues. (If you can't solve the problems, report them and continue with the assignment).**

```{r hyperoptimization_Barak_Logistic_Regression, echo=TRUE, message=FALSE}
# bayes_opt_25_rBayes.R
# Bayesian hyper‑parameter optimisation (25 evaluations) using **rBayesianOptimization**
# instead of ParBayesianOptimization to avoid the inBounds NA bug.

# 0. Install / load packages ---------------------------------------------------
required_pkgs <- c("brms", "loo", "rBayesianOptimization", "progress")
for (pkg in required_pkgs) if (!requireNamespace(pkg, quietly = TRUE)) install.packages(pkg)

library(brms)
library(loo)
library(rBayesianOptimization)   # <‑‑ different optimiser
library(progress)

# 1. Search space --------------------------------------------------------------
search_bounds <- list(
  sigma_fixed = c(0.5, 10),
  sigma_group = c(0.5, 5),
  adapt_delta = c(0.90, 0.99),
  iter        = c(1000, 2000)
)

# helper to snap iter to 1000 / 1500 / 2000
snap_iter <- function(x) {
  cuts <- c(1000, 1500, 2000)
  cuts[which.min(abs(cuts - x))]
}

# 2. Objective function --------------------------------------------------------
obj_fun <- function(sigma_fixed, sigma_group, adapt_delta, iter) {
  iter   <- snap_iter(iter)
  warmup <- iter / 2

  priors <- c(
    prior_string(sprintf("normal(0, %.4f)", sigma_fixed), class = "b"),
    prior_string(sprintf("student_t(3, 0, %.4f)", sigma_group), class = "sd")
  )

  score <- tryCatch({
    fit <- brm(
      Binary.Rating ~ Current.Ratio +
                     `Long.term.Debt...Capital` +
                     Debt.Equity.Ratio +
                     EBITDA.Margin +
                     `Operating.Cash.Flow.Per.Share` +
                     (1 | Sector),
      data    = train_binary,
      family  = bernoulli(link = "logit"),
      prior   = priors,
      iter    = iter,
      warmup  = warmup,
      control = list(adapt_delta = adapt_delta, max_treedepth = 12),
      cores   = 4,
      refresh = 0
    )
    -loo(fit, cores = 2)$estimates["looic", "Estimate"]  # maximise
  }, error = function(e) {
    message("Penalty – ", e$message)
    -1e6
  })

  list(Score = score)
}

# 3. Run Bayesian optimisation -------------------------------------------------
set.seed(123)
opt <- BayesianOptimization(
  FUN = obj_fun,
  bounds = search_bounds,
  init_points = 10,
  n_iter      = 15,
  acq = "ucb", kappa = 2.576,
  verbose = TRUE
)

print(opt$Best_Par)
```
```{r hyperparamater_saving, echo=TRUE, message=FALSE}

# 1) Build (and create) your output directory
hyper.out.dir <- file.path(wd, "Hyperparameter_Optimization")
if (!dir.exists(hyper.out.dir)) {
  dir.create(hyper.out.dir, recursive = TRUE)
}

# 2) Construct file paths
stamp <- format(Sys.time(), "%Y%m%d_%H%M%S")

csv_file <- file.path(
  hyper.out.dir,
  paste0("best_params_", stamp, ".csv")
)
rds_file <- file.path(
  hyper.out.dir,
  paste0("rBayesOpt_res_", stamp, ".rds")
)

# 3) Write the best‐params CSV
write.csv(
  opt$Best_Par,   # the best‐parameter data.frame
  file = csv_file,
  row.names = FALSE
)

# 4) Save the full optimization result as an RDS
saveRDS(
  opt,           # the full ParBayesianOptimization result object
  file = rds_file
)

```


<!-- message = FALSE, results = "hide" prevents displaying output, if you need to show something create another chunk of code -->

{r, message = FALSE, results = "hide"}
# models go here

```{r Barak_Logistic_Bayesian_Regression, echo=TRUE, message=FALSE}
# final_model_fit.R
# Fit a Bayesian multilevel logistic regression using the best hyperparameters
# (reads them directly from the uploaded CSV)

# 1. Load libraries (will error if not installed) ----------------------------
library(brms)
library(pROC)

# ---------------------------------------------------------------------------
# 2. Locate the best-params CSV ---------------------------------------------
# ---------------------------------------------------------------------------
# Assuming your working directory is stored in the variable `wd`.
# The CSV is expected in: <wd>/Hyperparameter_Optimization/

if (!exists("wd")) {
  wd <- getwd()  # fallback – uses current working directory
  message(sprintf("`wd` not found; defaulting to getwd(): %s", wd))
}

hyper.out.dir <- file.path(wd, "Hyperparameter_Optimization")
if (!dir.exists(hyper.out.dir)) {
  stop(sprintf("Directory '%s' does not exist – check `wd`.", hyper.out.dir))
}

csv_files <- list.files(hyper.out.dir, pattern = "^best_params_.*\\.csv$", full.names = TRUE)
if (length(csv_files) == 0) {
  stop("No 'best_params_*.csv' file found in ", hyper.out.dir)
}

# Pick the newest file (in case multiple exist)
csv_file <- csv_files[order(file.info(csv_files)$mtime, decreasing = TRUE)][1]
message(sprintf("Reading hyperparameters from: %s", basename(csv_file)))

best_params_df <- read.csv(csv_file, stringsAsFactors = FALSE)

# ---------------------------------------------------------------------------
# 3. Extract hyperparameters with sane fallbacks ----------------------------
# ---------------------------------------------------------------------------
get_or <- function(df, name, default) {
  if (name %in% names(df)) df[[name]][1] else default
}

sigma_fixed <- get_or(best_params_df, "sigma_fixed", 2.5)
sigma_group <- get_or(best_params_df, "sigma_group", 1.5)
adapt_delta <- get_or(best_params_df, "adapt_delta", 0.95)
iter        <- get_or(best_params_df, "iter", 1500)
warmup      <- floor(iter / 2)

# ---------------------------------------------------------------------------
# 4. Define priors -----------------------------------------------------------
# ---------------------------------------------------------------------------
priors <- c(
  prior_string(sprintf("normal(0, %.4f)", sigma_fixed), class = "b"),
  prior_string(sprintf("student_t(3, 0, %.4f)", sigma_group), class = "sd")
)

# ---------------------------------------------------------------------------
# 5. Fit model on `train_binary` ---------------------------------------------
# ---------------------------------------------------------------------------
fit <- brm(
  formula = bf(
    Binary.Rating ~ Current.Ratio +
                   `Long.term.Debt...Capital` +
                   Debt.Equity.Ratio +
                   EBITDA.Margin +
                   `Operating.Cash.Flow.Per.Share` +
                   (1 | Sector)
  ),
  data    = train_binary,
  family  = bernoulli(link = "logit"),
  prior   = priors,
  iter    = iter,
  warmup  = warmup,
  control = list(adapt_delta = adapt_delta, max_treedepth = 12),
  cores   = 4,
  refresh = 100
)

# ---------------------------------------------------------------------------
# 6. Evaluate on `test_binary` ----------------------------------------------
# ---------------------------------------------------------------------------
probs <- colMeans(posterior_linpred(fit, newdata = test_binary, transform = TRUE))

pred_class <- factor(ifelse(probs > 0.5, 1, 0), levels = c(0, 1))
true_class <- test_binary$Binary.Rating

conf_mat <- table(Predicted = pred_class, Actual = true_class)
accuracy  <- sum(diag(conf_mat)) / sum(conf_mat)
precision <- conf_mat["1", "1"] / sum(conf_mat["1", ])
recall    <- conf_mat["1", "1"] / sum(conf_mat[, "1"])
f1_score  <- 2 * precision * recall / (precision + recall)

roc_obj   <- roc(as.numeric(as.character(true_class)), probs)
auc_value <- auc(roc_obj)

# ---------------------------------------------------------------------------
# 7. Report ------------------------------------------------------------------
# ---------------------------------------------------------------------------
print(conf_mat)


# Build a small data‑frame and round only the numeric column -----------------
metrics <- data.frame(
  Metric = c("Accuracy", "Precision", "Recall", "F1 Score", "ROC AUC"),
  Value  = c(accuracy, precision, recall, f1_score, auc_value)
)
metrics$Value <- round(metrics$Value, 3)
print(metrics, row.names = FALSE)

# In case you still want individual lines in the console (use cat so they
# always show up in stdout)
cat(sprintf("\nAccuracy : %.3f\n", accuracy))
cat(sprintf("Precision: %.3f\n", precision))
cat(sprintf("Recall   : %.3f\n", recall))
cat(sprintf("F1 Score : %.3f\n", f1_score))
cat(sprintf("ROC AUC  : %.3f\n\n", auc_value))

# basic model diagnostics
print(fit)            # quick coefficients & R-hat
summary(fit)          # full convergence stats
bayesplot::mcmc_trace(as.array(fit))   # trace plots if you have bayesplot installed

# posterior predictive check
pp_check(fit)

# loo object for more detail (elpd, Pareto-k, etc.)
l <- loo(fit, cores = 2)
print(l)

```



```{r hyperoptimization_Jacob_Ordinal_Model, echo=TRUE, message=FALSE}
# parbayes_opt.R
# R script for Bayesian hyperparameter optimization using ParBayesianOptimization

# 0. Install and load packages
if (!requireNamespace("ParBayesianOptimization", quietly = TRUE)) {
  install.packages("ParBayesianOptimization", repos = 'https://cran.rstudio.com/')
}
if (!requireNamespace("brms", quietly = TRUE)) {
  install.packages("brms", repos = 'https://cran.rstudio.com/')
}
if (!requireNamespace("loo", quietly = TRUE)) {
  install.packages("loo", repos = 'https://cran.rstudio.com/')
}
library(ParBayesianOptimization)
library(brms)
library(loo)

# 1. Prepare training data
# Ensure 'train_binary' exists (this part is for the original bernoulli model, can be left)
if (!exists("train_binary")) {
  train_binary <- read.csv("Input_Data/df_binary.csv")
  train_binary$Binary.Rating <- factor(train_binary$Binary.Rating)
}

# --- BEGIN MODIFIED BRM MODEL CALL ---
# Define priors for the cumulative model with cloglog link and random intercepts
cloglog_intercept_priors <- c(
  prior(normal(0, 1.5), class = b),  # For all fixed effects, including interaction
  prior(student_t(3, 0, 2.5), class = sd), # For standard deviation of random intercepts
  prior(student_t(3, 0, 2.5), class = Intercept) # For cutpoints in cumulative model
)

# Ensure train_rating is available and Rating is an ordered factor
# It should be from the splitdata chunk, but we make sure Rating is ordered here.
if (!exists("train_rating")) {
    stop("train_rating dataframe not found. Ensure the 'splitdata' chunk has been run.")
}
rating_levels <- c("AAA","AA+","AA","AA-","A+","A","A-",
                   "BBB+","BBB","BBB-","BB+","BB","BB-",
                   "B+","B","B-","CCC","CC","C","D")
train_rating$Rating <- factor(train_rating$Rating, levels = rating_levels, ordered = TRUE)


# Fit a standalone brms model (cumulative with cloglog link and random intercepts only)
print("Fitting a cumulative brm model with cloglog link and random intercepts...")
fit_cumulative_cloglog_intercepts <- brm(
  formula = bf(
    Rating ~ Current.Ratio * Debt.Equity.Ratio + # Includes Current.Ratio, Debt.Equity.Ratio, and their interaction
             `Long.term.Debt...Capital` +
             EBITDA.Margin +
             `Operating.Cash.Flow.Per.Share` +
             (1 | Sector) # Random intercepts for Sector
  ),
  data    = train_rating,
  family  = cumulative(link = "cloglog"), # Cumulative likelihood with cloglog link
  prior   = cloglog_intercept_priors,
  iter    = 2000,
  warmup  = 1000,
  control = list(adapt_delta = 0.95),
  cores   = 4,
  refresh = 0,
  seed    = 123
)
print("Cumulative cloglog brm model fitting complete.")
print(summary(fit_cumulative_cloglog_intercepts))
# --- END MODIFIED BRM MODEL CALL ---

# 2. Define the objective function with error handling
# Returns a list with 'Score' to maximize (we use -LOOIC)
# obj_fun <- function(sigma_fixed, sigma_group, adapt_delta) {
#   result <- tryCatch({
#     # Build dynamic priors as strings
#     priors <- c(
#       prior_string(paste0("normal(0, ", round(sigma_fixed,4), ")"), class = "b"),
#       prior_string(paste0("student_t(3, 0, ", round(sigma_group,4), ")"), class = "sd")
#     )
# 
#     # Fit the brms model
#     fit <- brm(
#       formula = bf(
#         Binary.Rating ~ Current.Ratio +
#                         `Long.term.Debt...Capital` +
#                         Debt.Equity.Ratio +
#                         EBITDA.Margin +
#                         `Operating.Cash.Flow.Per.Share` +
#                         (1 | Sector)
#       ),
#       data    = train_binary,
#       family  = bernoulli(link = "logit"),
#       prior   = priors,
#       iter    = 1000,
#       warmup  = 500,
#       control = list(adapt_delta = adapt_delta),
#       cores   = 4,
#       refresh = 0
#     )
# 
#     # Compute LOOIC
#     looic_val <- loo(fit)$estimates["looic", "Estimate"]
#     list(Score = -looic_val)
#   }, error = function(e) {
#     message("Objective error: ", e$message)
#     return(NULL)
#   })
#   return(result)
# }

# 3. Define search bounds
# bounds <- list(
#   sigma_fixed = c(0.1, 10),
#   sigma_group = c(0.1, 5),
#   adapt_delta = c(0.8, 0.99)
# )

# 4. Run Bayesian Optimization with progress reporting
# set.seed(123)
# opt_res <- bayesOpt(
#   FUN           = obj_fun,
#   bounds        = bounds,
#   initPoints    = 5,
#   iters.n       = 15,
#   acq           = "ucb",       # Upper confidence bound
#   kappa         = 2.576,         # exploration/exploitation trade-off
#   verbose       = 2,             # textual progress
#   plotProgress  = TRUE,          # acquisition plot
#   dropNull      = TRUE           # skip failed evaluations
# )

# 5. Examine best hyperparameters
# best_params <- getBestPars(opt_res)
# print(best_params)
```

```{r Lilyanish_Ordinal_Model}

# Load required package for splitting
# (dplyr already loaded)

set.seed(123) # for reproducibility

# Split df_binary into 80% train and 20% test
n_binary <- nrow(df_binary)
train_idx_binary <- sample(seq_len(n_binary), size = 0.8 * n_binary)
train_binary <- df_binary[train_idx_binary, ]
test_binary <- df_binary[-train_idx_binary, ]

# Ensure factor levels of Binary.Rating are consistent
train_binary$Binary.Rating <- factor(train_binary$Binary.Rating)
test_binary$Binary.Rating <- factor(test_binary$Binary.Rating, levels = levels(train_binary$Binary.Rating))

# Split df_rating into 80% train and 20% test

n_rating <- nrow(df_rating)
train_idx_rating <- sample(seq_len(n_rating), size = 0.8 * n_rating)
train_rating <- df_rating[train_idx_rating, ]
test_rating <- df_rating[-train_idx_rating, ]

# Ensure factor levels of Rating are consistent
train_rating$Rating <- factor(train_rating$Rating)
test_rating$Rating <- factor(test_rating$Rating, levels = levels(train_rating$Rating))

#standardizing predictors so that priors later are easier to set 
predictorers <- c("Current.Ratio",
                  "Long.term.Debt...Capital",
                  "Debt.Equity.Ratio",
                  "EBITDA.Margin",
                  "Operating.Cash.Flow.Per.Share")
#Lilyana codee
library(brms)

model_l1 <- brm(
  formula = Rating ~ Debt.Equity.Ratio + Current.Ratio + 
    EBITDA.Margin + Operating.Cash.Flow.Per.Share + Long.term.Debt...Capital +
    ( 1 | Sector),
  data = train_rating,
  family = cumulative("logit"),
  seed = 123
)
summary(model_l1)
##All Rhat values = 1.00 => convergence
## Bulk and tail ESS >1000 & no warnings 
#pp checks lookedd good too 
#pp_check(model_l1)
#pp_check(model_l1, type="stat_2d")
#plot(model_l1) #nice fuzzy plots
```

```{r Jacob_Regression_Model}
# Best performing
library(brms)

train_binary <- janitor::clean_names(train_binary)

# Model formula
formula <- bf(binary_rating ~ debt_equity_ratio + current_ratio +
   roe_return_on_equity + net_profit_margin + 
   roa_return_on_assets + (1 | sector))

# Priors (informed from data exploration)
priors <- c(
  set_prior("normal(-0.5, 0.25)", class = "b", coef = "debt_equity_ratio"),
  set_prior("normal(0.2, 0.3)", class = "b", coef = "current_ratio"),
  set_prior("normal(0.4, 0.2)", class = "b", coef = "roe_return_on_equity"),
  set_prior("normal(0.25, 0.2)", class = "b", coef = "net_profit_margin"),
  set_prior("normal(0.5, 0.2)", class = "b", coef = "roa_return_on_assets"),
  set_prior("normal(0, 2)", class = "Intercept"))


# Fit final model using 6 chains and 4000 iterations
final_model_chip <- brm(
  formula = formula,
  data = train_binary,
  family = bernoulli(),
  chains = 6,
  iter = 4000,
  seed = 123,
  cores = 6,
)

```


```{r}
# Best performing
library(brms)

train_binary <- janitor::clean_names(train_binary)

# Model formula
formula <- bf(binary_rating ~ debt_equity_ratio + current_ratio +
   roe_return_on_equity + net_profit_margin + 
   roa_return_on_assets + (1 | sector))

# Priors (informed from data exploration)
priors <- c(
  set_prior("normal(-0.5, 0.25)", class = "b", coef = "debt_equity_ratio"),
  set_prior("normal(0.2, 0.3)", class = "b", coef = "current_ratio"),
  set_prior("normal(0.4, 0.2)", class = "b", coef = "roe_return_on_equity"),
  set_prior("normal(0.25, 0.2)", class = "b", coef = "net_profit_margin"),
  set_prior("normal(0.5, 0.2)", class = "b", coef = "roa_return_on_assets"),
  set_prior("normal(0, 2)", class = "Intercept"))


# Fit final model using 6 chains and 4000 iterations
final_model_chip <- brm(
  formula = formula,
  data = train_binary,
  family = bernoulli(),
  chains = 6,
  iter = 4000,
  seed = 123,
  cores = 6,
)

```

```{r}
summary(final_model_chip)

# For max Rhat:
max_rhat <- max(rhat(final_model_chip))
print(paste("Max Rhat:", max_rhat))

# For ESS:
min_bulk_ess <- min(neff_ratio(final_model_chip, type = "bulk"))
min_tail_ess <- min(neff_ratio(final_model_chip, type = "tail"))
print(paste("Min Bulk ESS:", min_bulk_ess))
print(paste("Min Tail ESS:", min_tail_ess))

# For divergent transitions (in Stan fit object):
sum(final_model_chip$fit@sim$diagnostics$divergent__)

# The model converged well, with Rhat values close to 1 and generally high effective sample sizes. Traceplots indicate good mixing. The fixed effects show mostly weak to moderate associations with the outcome. However, the calculated accuracy is currently 0%, likely due to misclassification when converting predicted probabilities to classes. After proper thresholding, this metric should be recalculated. We recommend conducting posterior predictive checks and verifying classification thresholds for more reliable evaluation.
```

b. Explain each model and describe its structure (what they assume about potential population-level or group-level effects), and the type of priors used. 

<!-- EXPLAIN BELOW, REFER TO EACH MODEL -->

## 4. Model checking (3pt)

a. Perform a prior sensitivity analysis for each model and modify the model if appropriate. Justify.

```{r Barak_Logistic_Bayesian_Regression_Sensitivity_Analysis, echo=TRUE, message=FALSE, warning=FALSE}
library(brms)
library(bayesplot)
library(ggplot2)

# 1. Load best hyperparameters from the most recent CSV
hyper.out.dir <- file.path(wd, "Hyperparameter_Optimization")
if (!dir.exists(hyper.out.dir)) {
  stop(sprintf("Directory '%s' does not exist – check `wd`.", hyper.out.dir))
}
csv_files <- list.files(hyper.out.dir, pattern = "^best_params_.*\\.csv$", full.names = TRUE)
if (length(csv_files) == 0) {
  stop("No 'best_params_*.csv' file found in ", hyper.out.dir)
}
csv_file <- csv_files[order(file.info(csv_files)$mtime, decreasing = TRUE)][1]
message(sprintf("Reading hyperparameters from: %s", basename(csv_file)))
best_params_df <- read.csv(csv_file, stringsAsFactors = FALSE)

# 2. Extract hyperparameters (with defaults as fallback)
get_or <- function(df, name, default) {
  if (name %in% names(df)) df[[name]][1] else default
}
sigma_fixed <- get_or(best_params_df, "sigma_fixed", 2.5)
sigma_group <- get_or(best_params_df, "sigma_group", 1.5)
adapt_delta <- get_or(best_params_df, "adapt_delta", 0.95)
iter        <- get_or(best_params_df, "iter", 1500)
warmup      <- floor(iter / 2)

# 3. Define original and alternative priors
orig_priors <- c(
  prior_string(sprintf("normal(0, %.4f)", sigma_fixed), class = "b"),
  prior_string(sprintf("student_t(3, 0, %.4f)", sigma_group), class = "sd")
)
alt_priors_tight <- c(
  prior_string("normal(0, 1)", class = "b"),
  prior_string(sprintf("student_t(3, 0, %.4f)", sigma_group), class = "sd")
)
alt_priors_wide <- c(
  prior_string("normal(0, 5)", class = "b"),
  prior_string(sprintf("student_t(3, 0, %.4f)", sigma_group), class = "sd")
)

# 4. Fit prior predictive (sample_prior = "only") under each prior set
pp_orig <- brm(
  formula      = Binary.Rating ~ Current.Ratio +
                                 `Long.term.Debt...Capital` +
                                 Debt.Equity.Ratio +
                                 EBITDA.Margin +
                                 `Operating.Cash.Flow.Per.Share` +
                                 (1 | Sector),
  data         = train_binary,
  family       = bernoulli(link = "logit"),
  prior        = orig_priors,
  sample_prior = "only",
  iter         = iter,
  warmup       = warmup,
  control      = list(adapt_delta = adapt_delta),
  chains       = 4,
  cores        = 4,
  refresh      = 0
)
pp_tight <- update(pp_orig, prior = alt_priors_tight, sample_prior = "only", refresh = 0)
pp_wide  <- update(pp_orig, prior = alt_priors_wide,  sample_prior = "only", refresh = 0)

# 5. Extract linear predictor draws, assign names, and plot
lin_orig_mat  <- as.matrix(posterior_linpred(pp_orig,  transform = FALSE))
colnames(lin_orig_mat)  <- paste0("obs", seq_len(ncol(lin_orig_mat)))
lin_tight_mat <- as.matrix(posterior_linpred(pp_tight, transform = FALSE))
colnames(lin_tight_mat) <- paste0("obs", seq_len(ncol(lin_tight_mat)))
lin_wide_mat  <- as.matrix(posterior_linpred(pp_wide,  transform = FALSE))
colnames(lin_wide_mat)  <- paste0("obs", seq_len(ncol(lin_wide_mat)))

mcmc_areas(
  lin_orig_mat[, 1:100],
  prob = 0.5
) + labs(title = "Original Priors")

mcmc_areas(
  lin_tight_mat[, 1:100],
  prob = 0.5
) + labs(title = "Tighter Priors (Normal(0,1))")

mcmc_areas(
  lin_wide_mat[, 1:100],
  prob = 0.5
) + labs(title = "Wider Priors (Normal(0,5))")

# 6. Re-fit the original model for posterior predictive check
fit_full <- brm(
  formula = Binary.Rating ~ Current.Ratio +
                           `Long.term.Debt...Capital` +
                           Debt.Equity.Ratio +
                           EBITDA.Margin +
                           `Operating.Cash.Flow.Per.Share` +
                           (1 | Sector),
  data    = train_binary,
  family  = bernoulli(link = "logit"),
  prior   = orig_priors,
  iter    = iter,
  warmup  = warmup,
  control = list(adapt_delta = adapt_delta, max_treedepth = 12),
  chains  = 4,
  cores   = 4,
  refresh = 0
)

pp_check(fit_full, ndraws = 100) + labs(title = "Posterior Predictive Check – Original Model")

```



```{r Lilyanish_Ordinal_Model_Sensitivity_Analysis}
#Prior sensitivity analysis and adjustment of priors 
#install.packages("priorsense")
library(priorsense)

sensitivity_l1 <- powerscale_sensitivity(model_l1)
print(sensitivity_l1)
#warnings for prior of intercept: potential prior-data conflict
#setting priors now
#Gelman et al 2020
#calculating sds for predictors:
sd_DER <- sd(train_rating$'Debt.Equity.Ratio', na.rm = TRUE)
sd_CR <- sd(train_rating$'Current.Ratio', na.rm = TRUE)
sd_EBITDA<- sd(train_rating$'EBITDA.Margin', na.rm = TRUE)
sd_OCF <- sd(train_rating$'Operating.Cash.Flow.Per.Share', na.rm = TRUE)
sd_LTDC <- sd(train_rating$'Long.term.Debt...Capital', na.rm = TRUE)

rating_numeric <-as.numeric(train_rating$Rating)

sd_R <-sd(rating_numeric, na.rm=TRUE)

prior_sd_DER <- 2.5*sd_R/sd_DER
#0.1631
prior_sd_CR <-2.5*sd_R/sd_CR
#6.0350
prior_sd_EBITDA <- 2.5*sd_R/sd_EBITDA
#0.2127
prior_sd_OCF <- 2.5*sd_R/sd_OCF
#1,4442
prior_sd_LTDC <- 2.5*sd_R/sd_LTDC
#32.2953

prios<-c(prior_sd_DER,prior_sd_CR, prior_sd_EBITDA, prior_sd_OCF, prior_sd_LTDC)
print(prios)


adj_priors<- c(
  prior(normal(0, 0.1631), class = "b", coef = "Debt.Equity.Ratio"),
  prior(normal(0, 6.0350), class = "b", coef = "Current.Ratio"),
  prior(normal(0, 0.2127), class = "b", coef = "EBITDA.Margin"),
  prior(normal(0, 1.4442), class = "b", coef = "Operating.Cash.Flow.Per.Share"),
  prior(normal(0, 32.2953), class = "b", coef = "Long.term.Debt...Capital"),
  prior(normal(0,10), class = "Intercept"),
  prior(exponential(1), class = "sd")
)

#new model
model_l2 <- brm(
  formula = Rating ~ Debt.Equity.Ratio + Current.Ratio + 
    EBITDA.Margin + Operating.Cash.Flow.Per.Share + Long.term.Debt...Capital +
    ( 1 | Sector),
  data = train_rating,
  family = cumulative("logit"),
  prior = adj_priors,
  seed = 123
)
summary(model_l2)
#Again Rhat and ESS are good, no warnings
#plot(model_l2) #nice and fuzzy
#pp_check(model_l2)
#pp_check(model_l2, type="stat_2d")
#pp_check(model_l2, type = "bars", ndraws = 100)
sensitivity_l2 <- powerscale_sensitivity(model_l2)
print(sensitivity_l2) #no problems priors are good

####This part idk where to put
#get posterior draws
pred_draws <- posterior_predict(model_l2)
#convert to binary where bbb-(ordinal index <= 11) and below are 0 (do not invest), above are 1 = invest
binary_draws <- pred_draws <= 11
#getting posterior mean probability of getting a binary rating = 1 (invest)
binary_probability <- colMeans(binary_draws)
#classifying using a 0.7 treshhold probability
binary_predict <- ifelse(binary_probability >= 0.7, 1, 0) #we want less false positives (so we don't get invest when we are not quite certain we want to invest)
actual_binary_rating <- train_rating$Binary.Rating

#confusion matrix: 
table(Predicted = binary_predict, Actual = actual_binary_rating)
print(table(Predicted = binary_predict, Actual = actual_binary_rating))
```

```{r}
# Chip's model
# Predict on test set (e.g., test_binary)
test_binary <- janitor::clean_names(test_binary)
preds <- predict(final_model_chip, newdata = test_binary, summary = TRUE)

```

```{r}
predicted_class <- ifelse(preds[, "Estimate"] >= 0.5, 1, 0)
true_class <- as.integer(test_binary$binary_rating)

accuracy <- mean(predicted_class == true_class)
print(paste("Accuracy:", accuracy * 100, "%"))

```

```{r}
accuracy <- mean(predicted_classes == true_classes)
print(paste("Accuracy:", round(accuracy * 100, 2), "%"))

# Confusion Matrix
table(Predicted = predicted_class, Actual = true_classes)

```

```{r}
install.packages("priorsense")
library(priorsense)
# 1. Define your original priors (same as before)
priors_informed <- c(
  set_prior("normal(-0.5, 0.25)", class = "b", coef = "debt_equity_ratio"),
  set_prior("normal(0.2, 0.3)", class = "b", coef = "current_ratio"),
  set_prior("normal(0.4, 0.2)", class = "b", coef = "roe_return_on_equity"),
  set_prior("normal(0.25, 0.2)", class = "b", coef = "net_profit_margin"),
  set_prior("normal(0.5, 0.2)", class = "b", coef = "roa_return_on_assets"),
  set_prior("normal(0, 2)", class = "Intercept")
)

# 2. Define your model formula
formula <- bf(binary_rating ~ debt_equity_ratio + current_ratio +
              roe_return_on_equity + net_profit_margin + 
              roa_return_on_assets)

# 3. Fit the model *with sample_prior = "yes"*
model_informed <- brm(
  formula = formula,
  data = train_binary,
  family = bernoulli(),
  prior = priors_informed,
  sample_prior = "yes",    # required for powerscale_sensitivity
  chains = 6,
  iter = 4000,
  seed = 123,
  cores = 6
)

# 4. Run the prior sensitivity analysis
ps_result <- powerscale_sensitivity(model_informed)

# 5. Plot sensitivity
plot(ps_result)
```

Chip's model: 
We conducted a prior sensitivity analysis using powerscale_sensitivity() and found that the posterior estimates were primarily driven by the likelihood rather than the prior. The diagnostic plot showed that the influence of the priors was minimal across parameters, with the majority of the posterior variability attributed to the data. This indicates that the model is robust to the chosen priors and that our conclusions are not overly sensitive to prior assumptions.



<!-- EXPLAIN CONCLUSIONS AND WHETHER MODELS ARE KEPT, MODIFIED  -->

b. Conduct posterior predictive checks for each model to assess how well they fit the data.
Explain what you conclude.

```{r}
# Chip's Model #
# posterior-predictive draws
yrep <- posterior_predict(final_model_chip, draws = 4000)

# convert outcome to numeric 0/1
y_obs <- as.numeric(as.character(train_binary$binary_rating))

# quick mean check
c(Observed = mean(y_obs), Predicted = mean(colMeans(yrep)))

# default bayesplot PPCs
ppc_bars(y_obs, yrep, prob = 0.9)
ppc_bars_grouped(y_obs, yrep, group = train_binary$sector, prob = 0.9)

```




```{r}

```

c. Binary Classification Metrics from Ordinal Model

While our primary model (`fit_cumulative_cloglog_intercepts`) predicts ordinal ratings, we can derive binary classification metrics by defining a threshold to group these ratings. For this exercise, we will classify ratings as either "Investment Grade" (AAA down to BBB-) or "Speculative Grade" (BB+ down to D).

We will predict probabilities for each rating category on the test set, sum the probabilities for all investment grade categories, and then use a 0.5 probability threshold on this sum to make a binary prediction.

```{r calculate-binary-metrics, echo=TRUE, message=FALSE, warning=FALSE}
# Ensure the brms model object is available (it should be if the hyperparameter chunk ran)
if (!exists("fit_cumulative_cloglog_intercepts")) {
  stop("Model object 'fit_cumulative_cloglog_intercepts' not found. Please ensure the model fitting chunk has been run.")
}

# Ensure the test_rating dataset is available
if (!exists("test_rating")) {
  stop("Test dataset 'test_rating' not found. Please ensure the data splitting chunk has been run.")
}

# Load necessary library for metrics if not already loaded
if (!requireNamespace("caret", quietly = TRUE)) {
  install.packages("caret", repos = 'https://cran.rstudio.com/')
}
library(caret)

# Define rating levels and investment grade threshold
rating_levels <- c("AAA","AA+","AA","AA-","A+","A","A-",
                   "BBB+","BBB","BBB-","BB+","BB","BB-",
                   "B+","B","B-","CCC","CC","C","D")
investment_grade_ratings <- c("AAA","AA+","AA","AA-","A+","A","A-", "BBB+","BBB","BBB-")

# Ensure Rating in test_rating is an ordered factor with correct levels
test_rating$Rating <- factor(test_rating$Rating, levels = rating_levels, ordered = TRUE)

# 1. Predict probabilities for each category on the test set
# posterior_epred gives expected probabilities for each category
# Using a smaller number of draws for faster execution during knitting, increase for more stable estimates.
pred_probs_ordinal <- posterior_epred(fit_cumulative_cloglog_intercepts, newdata = test_rating, ndraws = 500) 

# pred_probs_ordinal is an array: (draws, observations, categories)
# We want the mean probability for each category for each observation
mean_pred_probs_ordinal <- colMeans(pred_probs_ordinal) # Now (observations, categories)
if(ncol(mean_pred_probs_ordinal) == length(rating_levels)){
    colnames(mean_pred_probs_ordinal) <- rating_levels # Assign category names
} else {
    stop("Mismatch between number of predicted categories and length of rating_levels.")
}


# 2. Calculate the total probability of being investment grade for each observation
prob_investment_grade <- rowSums(mean_pred_probs_ordinal[, intersect(colnames(mean_pred_probs_ordinal), investment_grade_ratings), drop = FALSE])

# 3. Convert to binary predictions ("Investment" vs "Speculative") using a 0.5 threshold
predicted_binary_labels <- ifelse(prob_investment_grade >= 0.5, "Investment", "Speculative")
predicted_binary_labels <- factor(predicted_binary_labels, levels = c("Investment", "Speculative"))

# 4. Determine true binary labels from test_rating
true_binary_labels <- ifelse(test_rating$Rating %in% investment_grade_ratings, "Investment", "Speculative")
true_binary_labels <- factor(true_binary_labels, levels = c("Investment", "Speculative"))

# 5. Compute and print confusion matrix and metrics
# Check if there is variability in both true and predicted labels
if (length(unique(predicted_binary_labels)) < 2 || length(unique(true_binary_labels)) < 2) {
    message("Warning: One or both factors (predicted or true labels) have fewer than 2 levels (e.g., all predictions are 'Investment').")
    message("Standard confusion matrix metrics from caret might not be meaningful or directly computable in this scenario.")
    message("Predicted distribution:")
    print(table(predicted_binary_labels))
    message("True distribution:")
    print(table(true_binary_labels))
    # Add custom calculations if needed for this edge case, e.g. accuracy if all match.
    if(length(unique(predicted_binary_labels)) == 1 && length(unique(true_binary_labels)) == 1 && unique(predicted_binary_labels) == unique(true_binary_labels)){
      cat("Accuracy: 100% (all predictions match the single true class)\n")
    } else {
      cat("Accuracy, Precision, Recall, F1 cannot be reliably calculated by caret.\n")
    }
} else {
    conf_matrix_obj <- confusionMatrix(data = predicted_binary_labels, 
                                     reference = true_binary_labels, 
                                     positive = "Investment")

    print(conf_matrix_obj$table) # Print the confusion matrix table

    accuracy <- conf_matrix_obj$overall["Accuracy"]
    precision <- conf_matrix_obj$byClass["Precision"]
    recall <- conf_matrix_obj$byClass["Recall"]
    f1_score <- conf_matrix_obj$byClass["F1"]

    cat("\nBinary Classification Metrics (Positive Class: 'Investment'):\n")
    cat("Accuracy: ", ifelse(is.na(accuracy), "NA", round(accuracy, 4)), "\n")
    cat("Precision: ", ifelse(is.na(precision), "NA", round(precision, 4)), "\n")
    cat("Recall (Sensitivity): ", ifelse(is.na(recall), "NA", round(recall, 4)), "\n")
    cat("F1-Score: ", ifelse(is.na(f1_score), "NA", round(f1_score, 4)), "\n")
}
```



<!-- EXPLAIN CONCLUSIONS -->

## 5. Model Comparison (1.5pt)

a. Use k-fold cross-validation to compare the models.

```{r}

```
b. Determine the best model based on predictive accuracy and justify your decision.

<!-- DECISION -->


## 6. Interpretation of Important Parameters (1.5pt)

Choose one of the best models and interpret its most important parameters.

<!-- INTERPRETATION AND CODE GOES HERE -->
 


# Contributions of each member 

-
-
-
-

# Statement of technology


# References

<!-- Complete if necessary -->

